Abstractions and exemplars: the measure noun phrase alternation in German

### 2018-04-11 Latex/sections/3corpus.tex
Final author fixes. Off to proofreading.

For the present study, I used the German \textit{Corpus from the Web} (COW) in its 2014 version DECOW14A (\citealp{SchaeferBildhauer2012full}, and \citealp{Schaefer2015b}, as well as \citealp{BiemannEa2013}, and \citealp{SchaeferBildhauer2013}, for overviews of web corpora in general and the methodology of their construction), which contains almost 21 billion tokens.%
\footnote{The COW corpora (Dutch, English, French, German, Spanish, Swedish) are made available for free at \url{https://www.webcorpora.org}.
At the time of this writing, a newer 2016 version DECOW16A has already been released.}
I chose this corpus for two main reasons.
First, the external validity of any study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW14A corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo \citep{KupietzEa2010} of the Institute for the German Language (IDS), which contains almost exclusively newspaper texts.%
\footnote{It was shown in \cite{W16-2601} that, for example, the range of topics covered is much smaller in DeReKo compared to DECOW14A.}
Second, it was already mentioned that normative grammars often adopt clear positions regarding the grammaticality of either the \NACa\ or the \PGCa.
Thus, newspaper text or any other text that conforms strongly to normative grammars might not represent the alternation phenomenon fully (and without bias) because authors and proofreaders who must adhere to normative guidelines might favour one alternative or the other explicitly.
Web corpora, on the other hand, contain at least some amount of non-standard language from forums and similar sources.
For these or similar reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014}, \cite{VanGoethemHuening2015}, \cite{MuellerS2014}, \cite{Schaefer2016c}, \cite{SchaeferSayatz2014}, \cite{SchaeferSayatz2016}, and \cite{Zimmer2015}. 
Therefore, DECOW14A is a valid choice for this study.

@@

Among the factors potentially influencing the alternation (see Section~\ref{sec:germanmeasurenps}) are lemma-specific preference effects.
Therefore, it was desirable to obtain a sample in which most of the highly frequent actually-occurring combinations of kind nouns and measure nouns were represented.
I applied a two-stage process in order to obtain a list of co-occurring measure nouns and kind nouns.

In the first step, I exported a list of all nouns from the DECOW14A01 sub-corpus (one billion tokens large).
The one hundred most frequent mass nouns were extracted manually from this list.
Mass nouns were defined as concrete nouns which genuinely denote a substance, combine with uninflected mass quantifiers such as \textit{viel} `much' and \textit{wenig} `little' (\textit{viel Bier} `much beer'), and form only sortal and unit plurals (such as the plural \textit{Biere} `types of beer' or `glasses of beer').
Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are not quantified in the same way as concrete mass nouns.

This list of mass nouns was used in the second step to derive a list of measure nouns co-occurring with the mass nouns. 
In order to generate this list, I utilised the fact that a direct sequence of two nouns almost always instantiates the bare-noun NAC if the second noun is a mass noun.
I searched for all sequences N\Sub{1}N\Sub{2} where N\Sub{2} was one of the mass noun lemmas extracted in the first step.
Then, the resulting 100 lists of noun-noun combinations were each sorted by frequency in descending order and sieved manually to remove erroneous hits.
From each of the 100 lists, I also removed noun-noun combinations that had a frequency below 2, except if the individual list would have otherwise been shorter than 20 noun-noun combinations.
The result was a list of the most frequent 2,365 individual combinations of a measure noun and a mass noun.

@@

The variables \textit{Kindattraction} and \textit{Measureattraction} encode the ratio with which a given kind noun lemma or measure noun lemma occurs in the \PGCd\ and the \NACb.
They were calculated from auxiliary samples to be described in Section~\ref{sec:mainstudysampling} as a log-transformed quotient.
The higher the value for some noun, the (relatively) more often this noun occurs in the \PGCd.
It could be argued that other measures of attraction strength should be used, for example those popularised in collostructional analysis (CA; \citealp{StefanowitschGries2003,GriesStefanowitsch2004}; see also \citealp{Gries2015a}).
However, the goal here is to quantify how often lemmas occur in the \PGCd\ and the \NACb, and these constructions do not compete at all but are rather mutually exclusive.
While this does not preclude the use of collostructional analysis, it is an open question whether the marginals are cognitively relevant in this case.
After all, the main difference between the quotient used here and measures used in CA (originally signed logarithmised p-values from a Fisher test) is that they take the marginals into account. 
However, using Fisher-$p$-based collexeme strength instead of the raw frequency quotient was tried as an alternative (variables \textit{Kindcollo} and \textit{Measurecollo}); see Section~\ref{sec:mainstudystatisticalmodel} for a discussion of the negative results.

@@

Furthermore, to control for simple frequency effects, \textit{Kindfreq} and \textit{Measurefreq} are the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW14A corpus creators on their web page.

@@

The overall intercept is -5.370, and it represents \textit{Cardinal=Yes}, \textit{Measureclass=Physical} as 0 for all (z-transformed) numeric regressors.
The standard deviation in the random intercepts is 1.157 for \textit{Measurelemma} ($p_{\text{PB}}=0.001$) and 0.923 for \textit{Kindlemma} ($p_{\text{PB}}=0.002$).
Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.278$ and $R^2_c=0.566$ (see \citealp{Gries2015} or \citealp{NakagawaSchielzeth2013}).

@@

The coefficients of determination indicate that the effect of the fixed effects (marginal $R^2$) is adequately strong, and that lemma-specific effects are equally strong with a difference between the two coefficients of 0.289.

The main prototypical factors \textit{Cardinal} and \textit{Measureclass} both pass the PBmodcomp test at sig=0.05 and play the expected role.
Non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACb.
At the other end of the scale, nouns denoting natural portions like \textit{Haufen} `heap', \textit{Bündel} `bundle', \textit{Schluck} `gulp' favour the \PGCd.
Also, The presence of a cardinal modifier clearly favours the \NACb.
The stylistic factor \textit{Genitives} is confirmed to have an influence as predicted inasmuch as a general lack of genitives (which is encoded as a \textit{higher} value of the variable) favours the \NACb.
\textit{Badness} points into the wrong direction, but it barely reaches sig=0.05, and its confidence interval includes 0.
The lemma frequency regressors both clearly fail the PBmodcomp test.
I will return to these results in Section~\ref{sec:mainstudyinterpretation}.

@@

For the main study, each of the 2,365 noun–noun combinations was queried in the alternating constructions \PGCa\ and \NACa\ using a sub-corpus of DECOW14A (slices DECOW14A01 to DECOW14A10 with a total size of 10 billion tokens).
For further manual annotation, a random sub-sample was generated.
For each mass noun, the concordance was downsampled randomly to maximally 100 sentences, which resulted in 6,843 sentences.
In the manual annotation, the sample was further reduced to 5,063 sentences due to removal of noisy material, erroneous hits, and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.%
\footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were listed and queried, resulting in a sample of 871 sentences.
As stated in Section~\ref{sec:germanmeasurenps}, the \NACa\ is virtually never used with plural kind nouns, and this sample was not used except for quantifying the frequency of occurrence of the constructions (67 times \NACa\ and 794 times \PGCa).
The sample is distributed with the data package accompanying this paper.}

Finally, two auxiliary samples were also drawn.
As mentioned in Section~\ref{sec:analyses}, the distribution of the measure noun and kind noun lemmas in the \NACb\ and the \PGCd\ with a determiner will be modelled as factors influencing the alternation.
Therefore, all noun-noun pairs from the process described above were also queried in the two non-alternating constructions, resulting in 17,252 hits for the \PGCd\ and 315,635 hits for the \NACb.

@@

A multilevel logistic regression model was fit which models the influence of the regressors specified in Table~\ref{tab:variables} on the probability that the \PGCa\ is chosen over the \NACa.
All regressors from Table~\ref{tab:variables} were included, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
The sample size was \textit{n=}5,063 with 1,134 cases of \PGCa\ and 3,929 cases of \NACa.
The results of the estimation are shown in Table~\ref{tab:bigtable} and in Figure~\ref{fig:fixeffs}.
The intercept comprises \textit{Cardinal=Yes}, \textit{Measurecase=Nom}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}, and 0 for all numeric z-transformed regressors.
It was estimated at -3.548.

The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as group-level predictors (or second-level effects), cf.\ \citet[265--269,302--304]{GelmanHill2006}.
The same goes for those listed with the kind lemma as their unit of reference.
Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as favouring the \PGCa.

@@

Standard diagnostics show that the model quality is quite good.
Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.409$ and $R^2_c=0.495$.
The rate of correct predictions is 0.843, which means a proportional reduction of error of $\lambda=0.297$.
Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and the highest corrected $\text{GVIF}^{1/2\text{df}}$ was 1.520 for \textit{Cardinal}.
The lemma intercepts have standard deviations of $\sigma_{\text{Measurelemma}}=0.448$ and $\sigma_{\text{Kindlemma}}=0.604$.
Only \textit{Kindfreq} ($\mpPB=0.095$) could be seen as slightly too high to be convincing, failing at sig=0.05.

Using signed logarithmised Fisher p-values as a measure of collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) (see Section~\ref{sec:variablesandannotation}) was not successful.
While the attraction measures reach satisfying sig levels in the PBmodcomp test (0.020 for \textit{Kindattraction} and 0.001 for \textit{Measureattraction}), the $p_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.
The coefficients of determination drop to $R_m^2=0.376$ and $R^2_c=0.480$.

@@

The results reported in Section~\ref{sec:corpushierarchicalmodel} generally confirm the hypotheses from Section~\ref{sec:analyses}.
First, the prototypicality effect related to the non-alternating \PGCd\ and \NACb\ can be shown (see the effect plots in Figure~\ref{fig:eff:attraction}).%
\footnote{Effect plots were created using the \textit{effects} package \citep{Fox2003}.}
The effect is as expected:
if a lemma appears relatively more often in the \PGCd\ (compared to its frequency in the \NACb), the \PGCa\ tends to be chosen over the \NACa\ with this specific lemma.
The effect for measure nouns is stronger, and it was estimated with higher precision.

An interesting picture emerges for the lemma frequencies.
A higher-than-average lemma frequency of measure nouns favours the \NACa, which is as expected if we assume at least a tendency for highly grammaticalised items to be more frequent.
With kind nouns, higher frequency seems to favour the \PGCa.
However, this effect has no clear theoretical interpretation, and its coefficient estimate is imprecise (not significant at sig=0.05).
The effect can therefore be ignored or treated as a nuisance variable.

@@

Figure~\ref{fig:eff:leftcontext} shows that cardinals indeed influence the choice of the alternant, and that cardinals have a strong tendency to co-occur with the \NACa.
This effect was predicted in Section~\ref{sec:analyses}.

@@

The influence of \textit{Measurecase} is as predicted in previous analyses (see Section~\ref{sec:analyses}).
A measure noun in the dative favours the \PGCa\ (compared to the nominative, which is on the intercept).
Although \textit{Measurecase} is a nuisance variable in the context of this study, convergence with previous work strengthens its validity.

To close this section, I now compare the results of the pre-study and the main study.
Even though the non-alternating constructions are surely subject to additional constraints, the coefficients align neatly in many cases.
First of all, the intercepts encode a similar overall dominance of the NAC constructions (pre-study: -5.370; main study: -3.548).
For the \textit{Cardinal} effect, the coefficients 1.419 (pre-study) and 1.189 (main study) have the same sign and magnitude and mostly overlapping confidence intervals.
The levels of \textit{Measureclass} have comparable coefficients, although the divergence is larger.
In both studies, the non-referential measure nouns in the \textit{Physical} class (on the intercept) are most clearly associated with the NAC constructions.
Also, in both cases, the \textit{Container} class is closest to \textit{Physical} with the same sign and magnitude (pre-study: 0.445; main study: 0.252).
The main difference -- if we ignore the \textit{Rest} class which can be expected to show no clear tendency -- is that \textit{Amount} and \textit{Portion} are slightly closer together in their tendency to favour the PGC constructions in the pre-study (\textit{Amount} 1.597 and \textit{Portion} 1.782 in the pre-study vs.\ \textit{Amount} 0.831 and \textit{Portion} 1.217 in the main study).
The difference is not huge, and the overall order of the coefficients is the same.
The \textit{Genitives} effect also converges with -0.710 in the pre-study and -0.693 in the main study.
The two studies do not converge with respect to the \textit{Badness} variable, which is not significant in the pre-study.
However, it would be surprising if we achieved perfectly converging results given that even real effects are missed at certain rates in empirical studies.
In the case of \textit{Badness}, we see that even in the main study, the effect size is small (-0.152), and thus even at the impressive sample size of roughly 5,000 in both studies, the effect might simply be too weak to be detected reliably.
Finally, the frequency effect \textit{Measurefreq} is detected in the main study but not in the pre-study, while the \textit{Kindfreq} effect is essentially absent in both studies.
In connection with this, it is revealing to look at the coefficients of determination.
In the pre-study, a much greater proportion of the variance is explained by taking the lemma random effects into account ($R^2_m=0.278$ and $R^2_c=0.566$) than it is in the main study ($R^2_m=0.376$ and $R^2_c=0.480$).
Thus, the frequency effect for measure lemmas might be swamped by the lemma random intercepts.
All things considered, the studies have shown that the predicted prototype and exemplar effects are reflected in usage data for both the alternating and the non-alternating constructions.

>>>>

For the present study, I used the German \textit{Corpus from the Web} (COW) in its 2014 version DECOW14A (\citealp{SchaeferBildhauer2012full}, and \citealp{Schaefer2015b}, as well as \citealp{BiemannEa2013}, and \citealp{SchaeferBildhauer2013}, for overviews of web corpora in general and the methodology of their construction), which contains almost 21 billion tokens.%
\footnote{The COW corpora (Dutch, English, French, German, Spanish, Swedish) are made available for free at \url{https://www.webcorpora.org}.
At the time of this writing, a newer 2016 version DECOW16A has already been released.}
I chose this corpus for two main reasons.
First, the external validity of any study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW14A corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo of the Institute for the German Language \citep{KupietzEa2010}, which contains almost exclusively newspaper texts.%
\footnote{It was shown in \cite{W16-2601} that, for example, the range of topics is much smaller in DeReKo compared to DECOW14A.}
Second, it was already mentioned that normative grammars often adopt clear positions regarding the grammaticality of either the \NACa\ or the \PGCa.
Thus, newspaper text or any other text that conforms strongly to normative grammars might not represent the alternation phenomenon fully and without bias because authors and proofreaders who must adhere to normative guidelines might favour one alternative or the other explicitly.
Web corpora, on the other hand, contain a significant amount of non-standard language from forums and similar sources.
For such reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014}, \cite{VanGoethemHuening2015}, \cite{MuellerS2014}, \cite{Schaefer2016c}, \cite{SchaeferSayatz2014}, \cite{SchaeferSayatz2016}, and \cite{Zimmer2015}. 
Therefore, DECOW14A is a valid choice for this study.

@@

Among the factors potentially influencing the alternation (see Section~\ref{sec:germanmeasurenps}) are lemma-specific effects.
Therefore, it was desirable to obtain a sample in which most of the highly frequent actually-occurring combinations of kind nouns and measure nouns were represented.
I applied a two-stage process in order to obtain a list of co-occurring measure nouns and kind nouns.

In the first step, I exported a list of all nouns from the DECOW14A01 sub-corpus (one billion tokens large).
Then, the one hundred most frequent mass nouns were extracted manually from this list.
Mass nouns were defined as concrete nouns which genuinely denote a substance, combine with uninflected mass quantifiers such as \textit{viel} `much' and \textit{wenig} `little' (\textit{viel Bier} `much beer'), and form only sortal and unit plurals (such as the plural \textit{Biere} `types of beer' or `glasses of beer').
Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are not measured in the same way as concrete mass nouns.

This list of mass nouns was used in the second step to derive a list of measure nouns co-occurring with the mass nouns. 
In order to generate this list, I utilised the fact that a direct sequence of two nouns almost always instantiates the bare-noun NAC if the second noun is a mass noun.
I searched for all sequences N\Sub{1}N\Sub{2} where N\Sub{2} was one of the mass noun lemmas extracted in the first step.
Then, the resulting 100 lists of noun-noun combinations were each sorted by frequency in descending order and sieved manually to remove erroneous hits.
From each of the 100 lists, I also removed noun-noun combinations that had a frequency below 2, except if the individual list would have otherwise been shorter than 20 noun-noun combinations.
The result was a list of the most frequent 2,365 individual combinations of measure and mass nouns.

@@

The variables \textit{Kindattraction} and \textit{Measureattraction} encode the ratio with which a given kind noun lemma or measure noun lemma occurs in the \PGCd\ and the \NACb.
They were calculated from auxiliary samples to be described in Section~\ref{sec:mainstudysampling} as a log-transformed quotient.
The higher the value for some noun, the (relatively) more often this noun occurs in the \PGCd.
It could be argued that other measures of attraction strength should be used, for example those popularised in collostructional analysis (CA; \citealp{StefanowitschGries2003,GriesStefanowitsch2004}; see also \citealp{Gries2015a}).
However, the goal here is to quantify how often lemmas occur in the \PGCd\ and the \NACb, and these constructions do not compete at all but are rather mutually exclusive.
While this does not preclude the use of CA, it is an open question whether the marginals are cognitively relevant in this case.
After all, the main difference between the quotient used here and measures used in CA (originally signed logarithmised p-values from a Fisher test) is that they take the marginals into account. 
However, using Fisher-p-based collexeme strength instead of the raw frequency quotient was tried as an alternative (variables \textit{Kindcollo} and \textit{Measurecollo}).
See Section~\ref{sec:mainstudystatisticalmodel} for a discussion of the negative results.

@@

Furthermore, to control for simple frequency effects, \textit{Kindfreq} and \textit{Measurefreq} are included as the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW14A corpus creators on their web page.

@@

The overall intercept is -5.370, and it represents \textit{Cardinal=Yes}, \textit{Measureclass=Physical} and 0 for all (z-transformed) numeric regressors.
The standard deviation of the random intercepts is 1.157 for \textit{Measurelemma} ($p_{\text{PB}}=0.001$) and 0.923 for \textit{Kindlemma} ($p_{\text{PB}}=0.002$).
Nakagawa \& Schielzeth's pseudo-coefficients of determination are $R_m^2=0.278$ and $R^2_c=0.566$ (\citealp{Gries2015}; \citealp{NakagawaSchielzeth2013}).

@@

The coefficients of determination indicate that the influence of the fixed effects (marginal $R^2$) is adequately strong, and that lemma-specific effects are equally strong with a difference between the two coefficients of 0.289.

The main prototypical factors \textit{Cardinal} and \textit{Measureclass} both pass the PBmodcomp test at sig=0.05 and play the expected role.
Non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACb.
At the other end of the scale, nouns denoting natural portions like \textit{Haufen} `heap', \textit{Bündel} `bundle', \textit{Schluck} `gulp' favour the \PGCd.
Also, The presence of a cardinal modifier clearly favours the \NACb.
The stylistic factor \textit{Genitives} shows the predicted influence inasmuch as a general lack of genitives (which is encoded as a \textit{higher} value of the variable) favours the \NACb.
\textit{Badness} points into the wrong direction, but it barely reaches sig=0.05, and its confidence interval includes 0.
The lemma frequency regressors both clearly fail the PBmodcomp test.
I will return to these results in Section~\ref{sec:mainstudyinterpretation}.

@@

For the main study, each of the 2,365 noun–noun combinations was queried in the alternating constructions \PGCa\ and \NACa\ using a sub-corpus of DECOW14A (slices DECOW14A01 to DECOW14A10 with a total size of 10 billion tokens).
For further manual annotation, a random sub-sample was generated.
For each mass noun, the concordance was downsampled randomly to maximally 100 sentences, which resulted in 6,843 sentences.
In the manual annotation, the sample was further reduced to 5,063 sentences due to removal of noisy material, erroneous hits, and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.%
\footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were listed and queried, resulting in a sample of 871 sentences.
As stated in Section~\ref{sec:themeasurenounphrasealternation}, the \NACa\ is virtually never used with plural kind nouns, and this sample was not used except for quantifying the frequency of occurrence of the constructions (67 times \NACa\ and 794 times \PGCa).
The sample is distributed with the data package accompanying this paper.}

Finally, two auxiliary samples were also drawn.
As mentioned in Section~\ref{sec:analyses}, the distribution of the measure and kind noun lemmas in the \NACb\ and the \PGCd\ with a determiner will be modelled as factors influencing the alternation.
Therefore, all noun-noun pairs from the process described above were also queried in the two non-alternating constructions, resulting in 17,252 hits for the \PGCd\ and 315,635 hits for the \NACb.

@@

A multilevel logistic regression model was fit which models the influence of the regressors specified in Table~\ref{tab:variables} on the probability that the \PGCa\ is chosen over the \NACa.
All regressors from Table~\ref{tab:variables} were included, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
The sample size was \textit{n=}5,063 with 1,134 cases of \PGCa\ and 3,929 cases of \NACa.
The results of the estimation are shown in Figure~\ref{fig:fixeffs} and in Table~\ref{tab:bigtable}.
The intercept comprises \textit{Cardinal=Yes}, \textit{Measurecase=Nom}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}, and 0 for all numeric z-transformed regressors.
It was estimated at -3.548.

The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as group-level predictors (second-level effects), cf.\ \citet[265--269,302--304]{GelmanHill2006}.
The same goes for those listed with the kind lemma as their unit of reference.
Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as favouring the \PGCa.

@@

Standard diagnostics show that the model quality is quite good.
Nakagawa \& Schielzeth's pseudo-coefficients of determination are $R_m^2=0.409$ and $R^2_c=0.495$.
The rate of correct predictions is 0.843, which means a proportional reduction of error of $\lambda=0.297$.
Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and the highest corrected $\text{GVIF}^{1/2\text{df}}$ was 1.520 for \textit{Cardinal}.
The lemma intercepts have standard deviations of $\sigma_{\text{Measurelemma}}=0.448$ and $\sigma_{\text{Kindlemma}}=0.604$.
Only \textit{Kindfreq} ($\mpPB=0.095$) could be seen as slightly too high to be convincing, failing at sig=0.05.

Using signed logarithmised Fisher p-values as a measure of collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) (see Section~\ref{sec:variablesandannotation}) was not successful.
The $p_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.
The coefficients of determination dropped to $R_m^2=0.376$ and $R^2_c=0.480$.

@@

The results generally confirm the hypotheses from Section~\ref{sec:analyses}.
First, the exemplar effect (the influence of the non-alternating \PGCd\ and \NACb) was shown (see the effect plots in Figure~\ref{fig:eff:attraction}).%
\footnote{Effect plots were created using the \textit{effects} package \citep{Fox2003}.}
The effect is as expected:
if a lemma appears relatively more often in the \PGCd\ (compared to its frequency in the \NACb), the \PGCa\ tends to be chosen over the \NACa\ with this specific lemma.
The effect for measure nouns is stronger, and it was estimated with higher precision.

An interesting picture emerges for the lemma frequencies.
A higher-than-average lemma frequency of measure nouns favours the \NACa, which is as expected if we assume at least a tendency for highly grammaticalised words to be more frequent.
With kind nouns, higher frequency seems to favour the \PGCa.
However, this effect has no clear theoretical interpretation, and its coefficient estimate is imprecise (not significant at sig=0.05).
The effect can therefore be ignored or treated as a nuisance variable.

@@

Figure~\ref{fig:eff:leftcontext} shows that cardinals indeed influence the choice of the alternant, and that cardinals have a strong tendency to co-occur with the \NACa.

@@

The influence of \textit{Measurecase} is as predicted in previous analyses.
A measure noun in the dative favours the \PGCa\ (compared to the nominative, which is on the intercept).
Although \textit{Measurecase} is a nuisance variable in the context of this study, convergence with previous work strengthens its validity.

To close this section, I now compare the results of the pre-study and the main study.
Even though the non-alternating constructions are surely subject to additional constraints, the coefficients align neatly in many cases.
First of all, the intercepts encode a similar overall dominance of the NAC constructions (pre-study: -5.370; main study: -3.548).
For the \textit{Cardinal} effect, the coefficients 1.419 (pre-study) and 1.189 (main study) have the same sign and magnitude and mostly overlapping confidence intervals.
The levels of \textit{Measureclass} have comparable coefficients, although the divergence is larger.
In both studies, the non-referential measure nouns in the \textit{Physical} class (on the intercept) are most clearly associated with the NAC constructions.
Also, in both cases, the \textit{Container} class is closest to \textit{Physical} with the same sign and magnitude (pre-study: 0.445; main study: 0.252).
The main difference -- if we ignore the \textit{Rest} class which can be expected to show no clear tendency -- is that \textit{Amount} and \textit{Portion} are slightly closer together in their tendency to favour the PGC constructions in the pre-study (\textit{Amount} 1.597 and \textit{Portion} 1.782 in the pre-study vs.\ \textit{Amount} 0.831 and \textit{Portion} 1.217 in the main study).
The difference is not huge, and the overall order of the coefficients is the same.
The \textit{Genitives} effect also converges with -0.710 in the pre-study and -0.693 in the main study.
The two studies do not converge with respect to the \textit{Badness} variable, which is not significant in the pre-study.
However, it would be surprising if we achieved perfectly converging results given that even real effects are missed at certain rates in empirical studies.
In the case of \textit{Badness}, we see that even in the main study, the post-hoc effect size is small (-0.152), and thus even at the impressive sample size of roughly 5,000 in both studies, the effect might simply be too weak to be detected reliably.
Finally, the frequency effect \textit{Measurefreq} is detected in the main study but not in the pre-study, while the \textit{Kindfreq} effect is essentially absent in both studies.
In connection with this, it is revealing to look at the coefficients of determination.
In the pre-study, a much greater proportion of the variance is explained by taking the lemma random effects into account ($R^2_m=0.278$ and $R^2_c=0.566$) than it is in the main study ($R^2_m=0.376$ and $R^2_c=0.480$).
Thus, the frequency effect for measure lemmas might be swamped by the lemma random intercepts.
All things considered, the studies have shown that the predicted prototype and exemplar effects are reflected in usage data for both the alternating and the non-alternating constructions.
